---
title: "Practical Machine Learning Project"
author: "Alex Eller"
date: "January 22, 2016"
output:
  html_document:
    keep_md: yes
    fig_height: 5
    fig_width: 5
---
<style type="text/css">

body, td, h1, h2 {
   font-size: 12px;
}
code.r{
  font-size: 10px;
}
pre {
  font-size: 10px
}
</style>

### Overview
In this project, the goal is to use the data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict/quantify how well they do the exercise.

### Load the training and test data, and scrub the data
```{r,echo=TRUE, message=FALSE, warning=FALSE}
library(RCurl)
library(caret)
library(doParallel)

fileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
x <- getURL(fileUrl)
pml_training <- read.csv(text = x, na.strings = c("", "NA"))

fileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
x <- getURL(fileUrl)
pml_testing <- read.csv(text = x, na.strings = c("", "NA"))
```

Clean up the data by removing columns with nearly all/all NA's, and also remove the first columns as they are not relevant to the task at hand (id's, time data...)
```{r,echo=TRUE}
noNaTraining <- pml_training[,colSums(is.na(pml_training))<nrow(pml_training)*.97][8:60]
```

Find and remove columns that are significantly correlated and will not add any additional benefit to the model
```{r,echo=TRUE}
set.seed(8675309)
subset <- subset(noNaTraining, select=-c(classe))
correlations <- cor(subset)
highCorrelations <- findCorrelation(abs(correlations),0.90)
noNaTraining <- noNaTraining[,-highCorrelations]
```

### Partition training/testing datasets for measuring out of sample error rates.
```{r,echo=TRUE}
inTrain = createDataPartition(noNaTraining$classe, p = 0.75, list = FALSE)
training = noNaTraining[inTrain,]
testing = noNaTraining[-inTrain,]
```

### Prep for modeling, tune code for using multiple cores in parallel to speed up processing, also setup for 10 fold cross validation for training the model
```{r,echo=TRUE}
cluster <- makeCluster(detectCores() - 1) # leave 1 core for OS
registerDoParallel(cluster)
fitControl <- trainControl(method = "cv",
                           number = 10,
                           allowParallel = TRUE)
```

### Fit the model and stop the cluster to release the cores from service
```{r,echo=TRUE, message=FALSE, warning=FALSE, cache=TRUE}
modfit <- train(classe~.,data=training,method="rf",trControl = fitControl)
stopCluster(cluster)
```

Check the model
```{r,echo=TRUE}
modfit
```

Check the models stats/cross validation by viewing a confusionMatrix, the accuracy looks good
```{r,echo=TRUE}
confusionMatrix.train(modfit)
```

### Check the out of sample error by viewing a confusion matrix of the prediction
```{r,echo=TRUE, message=FALSE, warning=FALSE}
pred <- predict(modfit, testing)
confusionMatrix(pred, testing$classe)
```

### Run the prediction on the actual test set to generate the final results for the quiz answers
```{r,echo=TRUE}
pred <- predict(modfit, pml_testing)
pred
```
